# Resampling Methods

This lab will show us how to perform different resampling techniques. Some of these tasks are quite general and useful in many different areas. The bootstrap being such an example. This chapter introduces a lot of new packages.
This chapter will bring [rsample](https://www.tidymodels.org/start/resampling/) into view for creating resampled data frames as well as [yardstick](https://yardstick.tidymodels.org/) to calculate performance metrics. Lastly, will we also use [tune](https://tune.tidymodels.org/) to fit out models within said resamples. We also see a use of [dials](https://dials.tidymodels.org/) which are used together with tune to select hyperparameter tuning values.

```{r}
library(tidymodels)
library(ISLR)

Auto <- tibble(Auto)
```

## The Validation Set Approach

When fitting a model it is often desired to be able to calculate a performance metric to quantify how well the model fits the data. If a model is evaluated on the data it was fit on you are quite likely to get over-optimistic results. It is therefore we split our data into testing and training. This way we can fit the model to data and evaluate it on some other that that is similar.

Splitting of the data is done using random sampling, so it is advised to set a seed before splitting to assure we can reproduce the results.
The `inintial_split()` function takes a data.frame and returns a `rsplit` object. This object contains information about which observations belong to which data set, testing, and training. This is where you would normally set a proportion of data that is used for training and how much is used for evaluation. This is set using the `prop` argument which I set to `0.5` to closely match what happened in ISLR. I'm also setting the `strata` argument. This argument makes sure that both sides of the split have roughly the same distribution for each value of `strata`. If a numeric variable is passed to `strata` then it is binned and distributions are matched within bins.

```{r}
set.seed(1)
Auto_split <- initial_split(Auto, strata = mpg, prop = 0.5)
Auto_split
```

The testing and training data sets can be materialized using the `testing()` and `training()` functions respectively.

```{r}
Auto_train <- training(Auto_split)
Auto_test <- testing(Auto_split)
```

And by looking at `Auto_train` and `Auto_test` we see that the lengths match what we expect.

```{r}
Auto_train

Auto_test
```

Now that we have a train-test split let us fit some models and evaluate their performance. Before we move on it is important to reiterate that you should only use the testing data set once! Once you have looked at the performance on the testing data set you should not modify your models. If you do you might overfit the model due to data leakage.

Our modeling goal is to predict `mpg` by `horsepower` using a simple linear regression model, and a polynomial regression model. 
First, we set up a linear regression specification.

```{r}
lm_spec <- linear_reg() %>%
  set_mode("regression") %>%
  set_engine("lm")
```

And we fit it like normal. Note that we are fitting it using `Auto_train`.

```{r}
lm_fit <- lm_spec %>% 
  fit(mpg ~ horsepower, data = Auto_train)
```

We can now use the `augment()` function to extract the prediction and `rmse()` to calculate the root mean squared error. This will be the testing RMSE since we are evaluating on `Auto_test`.

```{r}
augment(lm_fit, new_data = Auto_test) %>%
  rmse(truth = mpg, estimate = .pred)
```

and we get a RMSE of `r augment(lm_fit, new_data = Auto_test) %>% rmse(truth = mpg, estimate = .pred) %>% pull(.estimate)`. This particular value is going to vary depending on what seed number you picked since the random sampling used in splitting the data set will be slightly different.

Using this framework makes it easy for us to calculate the training RMSE

```{r}
augment(lm_fit, new_data = Auto_train) %>%
  rmse(truth = mpg, estimate = .pred)
```

Comparing these two values can give us a look into how generalizable the model is to data it hasn't seen before. We do expect that the training RMSE to be lower than the testing RMSE but if you see a large difference there is an indication of overfitting or a shift between the training data set and testing data set. We don't expect a shift here since the data sets were created with random sampling.

Next we will fit a polynomial regression model. We can use the linear model specification `lm_spec` to add a preprocessing unit with `recipe()` and `step_poly()` to create the polynomial expansion of `horsepower`. we can combine these two with `workflow()` to create a workflow object.

```{r}
poly_rec <- recipe(mpg ~ horsepower, data = Auto_train) %>%
  step_poly(horsepower, degree = 2)

poly_wf <- workflow() %>%
  add_recipe(poly_rec) %>%
  add_model(lm_spec)

poly_wf
```

We can now fit this model. Again remember to fit it on the training data set `Auto_train`.

```{r}
poly_fit <- fit(poly_wf, data = Auto_train)
```

The testing RMSE is then calculated as

```{r}
augment(poly_fit, new_data = Auto_test) %>%
  rmse(truth = mpg, estimate = .pred)
```

Which is a little bit lower. So it would appear just from this, that the polynomial regression model has a better fit. Note that we are making decisions using the testing performance metrics, not the training performance metrics.

Lastly, we show below how changing the seed results in a slightly different estimate.

```{r}
set.seed(2)
Auto_split <- initial_split(Auto)

Auto_train <- training(Auto_split)
Auto_test <- testing(Auto_split)

poly_fit <- fit(poly_wf, data = Auto_train)

augment(poly_fit, new_data = Auto_test) %>%
  rmse(truth = mpg, estimate = .pred)
```

## Leave-One-Out Cross-Validation

Leave-One-Out Cross-Validation is not integrated into the broader tidymodels framework. For more information read [here](https://www.tmwr.org/resampling.html#leave-one-out-cross-validation).

## k-Fold Cross-Validation

Earlier we set `degree = 2` to create a second-degree polynomial regression model. But suppose we want to find the best value of `degree` that yields the "closest" fit. This is known as hyperparameter tuning and it is a case where we can use k-Fold Cross-Validation. To use k-Fold Cross-Validation we will be using the `tune` package, and we need 3 things to get it working:

- A parsnip/workflow object with one or more arguments marked for tuning,
- A `vfold_cv` rsample object of the cross-validation resamples,
- A tibble denoting the values of hyperparameter values to be explored.

we are doing the hyperparameter tuning on just one parameter, namely the `degree` argument in `step_poly()`. Creating a new recipe with `degree = tune()` indicated that we intend for `degree` to be tuned.

```{r}
poly_tuned_rec <- recipe(mpg ~ horsepower, data = Auto_train) %>%
  step_poly(horsepower, degree = tune())

poly_tuned_wf <- workflow() %>%
  add_recipe(poly_tuned_rec) %>%
  add_model(lm_spec)
```

This means that would not be able to fit this workflow right now as the value of `degree` is unspecified, and if we try we get an error:

```{r, warning=FALSE}
fit(poly_tuned_wf, data = Auto_train)
```

The next thing we need to create is the k-Fold data set. This can be done using the `vfold_cv()` function. Note that the function uses `v` instead of *k* which is the terminology of ISLR. we set `v = 10` as a common choice for *k*.

```{r}
Auto_folds <- vfold_cv(Auto_train, v = 10)
Auto_folds
```

The result is a tibble of `vfold_split`s which is quite similar to the `rsplit` object we saw earlier.

The last thing we need is a tibble of possible values we want to explore. Each of the tunable parameters in tidymodels has an associated function in the [dials package](https://dials.tidymodels.org/reference/index.html). We need to use the `degree()` function here, and we extend the range to have a max of 10. This dials function is then passed to `grid_regular()` to create a regular grid of values.

```{r}
degree_grid <- grid_regular(degree(range = c(1, 10)), levels = 10)
```

Using `grid_regular()` is a little overkill for this application since the following code would provide the same result. But once you have multiple parameters you want to tune it makes sure that everything is in check and properly named.

```{r}
degree_grid <- tibble(degree = seq(1, 10))
```

Now that all the necessary objects have been created we can pass them to `tune_grid()` which will fit the models within each fold for each value specified in `degree_grid`.

```{r}
tune_res <- tune_grid(
  object = poly_tuned_wf, 
  resamples = Auto_folds, 
  grid = degree_grid
)
```

It can be helpful to add `control = control_grid(verbose = TRUE)`, this will print out the progress. Especially helpful when the models take a while to fit. `tune_res` by itself isn't easily readable. Luckily `tune` provides a handful of helper functions.

`autoplot()` gives a visual overview of the performance of different hyperparameter pairs.

```{r}
autoplot(tune_res)
```

It appears that the biggest jump in performance comes from going to `degree = 2`. Afterward, there might be a little bit of improvement but it isn't as obvious.

The number used for plotting can be extracted directly with `collect_metrics()`. We also get an estimate of the standard error of the performance metric. We get this since we have 10 different estimates, one for each fold.

```{r}
collect_metrics(tune_res)
```

You can also use `show_best()` to only show the best performing models.

```{r}
show_best(tune_res, metric = "rmse")
```

We did see that the performance plateaued after `degree = 2`. There are a couple of function to select models by more sophisticated rules. `select_by_one_std_err()` and `select_by_pct_loss()`.  Here we use `select_by_one_std_err()` which selects the most simple model that is within one standard error of the numerically optimal results. We need to specify `degree` to tell `select_by_one_std_err()` which direction is more simple.

You want to

- use `desc(you_model_parameter)` if larger values lead to a simpler model
- use `you_model_parameter` if smaller values lead to a simpler model

lower polynomials models are simpler so we ditch `desc()`.

```{r}
select_by_one_std_err(tune_res, degree, metric = "rmse")
```

This selected `degree = 2`. And we will use this value since we simpler models sometimes can be very beneficial. Especially if we want to explain what happens in it.

```{r}
best_degree <- select_by_one_std_err(tune_res, degree, metric = "rmse")
```

This selected value can be now be used to specify the previous unspecified `degree` argument in `poly_wf` using `finalize_workflow()`.

```{r}
final_wf <- finalize_workflow(poly_wf, best_degree)

final_wf
```

This workflow can now be fitted. And we want to make sure we fit it on the full training data set.

```{r}
final_fit <- fit(final_wf, Auto_train)

final_fit
```

## The Bootstrap

```{r}
Portfolio_boots <- bootstraps(Portfolio, times = 1000)

alpha.fn <- function(split) {
  data <- analysis(split)
  X <- data$X
  Y <- data$Y
  
  (var(Y) - cov(X, Y)) / (var(X) + var(Y) - 2 * cov(X, Y))
}

alpha_res <- Portfolio_boots %>%
  mutate(alpha = map_dbl(splits, alpha.fn))

alpha_res
```

```{r}
Auto_boots <- bootstraps(Auto)

boot.fn <- function(split) {
  lm_fit <- lm_spec %>% fit(mpg ~ horsepower, data = analysis(split))
  tidy(lm_fit)
}

boot_res <- Auto_boots %>%
  mutate(models = map(splits, boot.fn))

boot_res %>%
  unnest(cols = c(models)) %>%
  group_by(term) %>%
  summarise(mean = mean(estimate),
            sd = sd(estimate))
```
